A formalization of semi-simplicial and semi-cubical sets

### 2023-01-10 paper/paper.tex
paper, tab-coh: use rotating to fix last table
paper: fix some underfull hboxes

\usepackage{multirow, float, tabularx, environ, booktabs, setspace, makecell, caption, art.cls/joinargs, colortbl, multicol, pdflscape, minted}
\usepackage[prefix=bonak]{art.cls/xkeymask}
\usepackage[skip=0.4em]{parskip}
\usetikzlibrary{patterns}

@@

\section{Overture}
In the context of typed functional programming, Reynolds' parametricity~\cite{reynolds72} interprets types as relations characterizing the observational equality of programs of this type. Parametricity can be iterated, and it has been noted that iterated Reynolds' parametricity has a cubical flavor~\cite{johann17,altenkirch15,moulin16}. We obtain a \emph{unary} variant of Reynolds' \emph{binary} parametricity by using predicates instead of relations, in which case we obtain a form of realizability~\cite{bernardy12,moulin16}. It has then been noted that iterated unary parametricity has an augmented simplicial flavor~\footnote{Observation by Hugo Moeneclaey}.

@@

Our precise mechanization in Coq is available at \href{https://github.com/artagnon/bonak}{github.com/artagnon/bonak}.

@@

While ordinary semi-simplicial sets are presheaves over the semi-simplex category $\hat{\Delta}$, augmented semi-simplicial sets are presheaves over, what we will define as $\DeltaPlus$. There are different ways to define $\DeltaPlus$, upto equivalence, and we use a combinatorial definition; the reason for our choice is that this definition has a straightforward extension to semi-cubical sets.

\begin{notation}[Sequences of integers]
  We write $\{i_1, \ldots i_n\}$ for sequences of integers. In particular, the empty sequence is written $\{\}$.
\end{notation}

@@

\begin{example}[$\Cube^1$]
  $\Cube^1$ is two points, given by $\Hom(0, 1)$ and a line, given by $\Hom(1, 1)$:

@@

\begin{example}[$\Cube^2$]
  $\Cube^2$ is four points, given by $\Hom(0, 2)$, four lines connecting the four points, given by $\Hom(1, 2)$, and a filler, given by $\Hom(2, 2)$:

@@

\begin{table}[H]
  \begin{tabularx}{\linewidth}{p{.2\linewidth}|p{.4\linewidth}|p{.3\linewidth}}
    \toprule
    Value of $\nu$ & 1                               & 2                  \\
    \graymidrule
    Interpretation & Augmented semi-simplicial types & Semi-cubical types \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

In category theory, the correspondence between fibered and indexed way, manifests as the correspondence between Grothendieck fibrations and pseudofunctors, of which a particular instance is the correspondence between discrete fibrations and presheaves, or even between discrete fibrations over sets and families of sets.

@@

$\restrl[n][p][\epsilon](d)(c_L, c_R)$ produces a layer, of which $\omega$-component has the type

@@

\subsection{Formal construction in extensional type theory}
% Abbreviated tables in this section
%
% The rule for implicit arguments for each table:
% show 1 explicit in first two parts
% show E + 1 explicit in the third part
\appendmask[bonak]{layer}[D]
\appendmask[bonak]{filler}[D]
\appendmask[bonak]{restrframe}[D]
\appendmask[bonak]{restrlayer}[D, d]
\appendmask[bonak]{restrfiller}[D, d]
\appendmask[bonak]{cohframe}[D]
\appendmask[bonak]{cohlayer}[D, d]
\appendmask[bonak]{cohfiller}[D, d]

@@

\subsection{Formal construction in intensional type theory\label{sec:itt}}
\renewcommandx{\cohf}{\coh{frame}[][][][][][][][]}

@@

\begin{landscape}
  \renewcommand*{\lab}{fulltab:coh}
  \input{tab-coh.tex}
\end{landscape}

>>>>

\usepackage{multirow, float, tabularx, environ, booktabs, setspace, makecell, caption, art.cls/joinargs, colortbl, multicol, minted, rotating}
\usepackage[prefix=bonak]{art.cls/xkeymask}
\usepackage[skip=0.4em]{parskip}
\usetikzlibrary{patterns}

@@

\section{Overture}
In the context of functional programming, Reynolds' parametricity~\cite{reynolds72} interprets types as relations characterizing the observational equality of programs of this type. Parametricity can be iterated, and it has been noted that iterated Reynolds' parametricity has a cubical flavor~\cite{johann17,altenkirch15,moulin16}. We obtain a \emph{unary} variant of Reynolds' \emph{binary} parametricity by using predicates instead of relations, in which case we obtain a form of realizability~\cite{bernardy12,moulin16}. It has then been noted that iterated unary parametricity has an augmented simplicial flavor~\footnote{Observation by Hugo Moeneclaey}.

@@

Our precise mechanization in Coq is available on Github, at \href{https://github.com/artagnon/bonak}{github.com/artagnon/bonak}.

@@

While ordinary semi-simplicial sets are presheaves over the semi-simplex category $\hat{\Delta}$, augmented semi-simplicial sets are presheaves over, what we will define as, $\DeltaPlus$. There are different ways to define $\DeltaPlus$, upto equivalence, and we use a combinatorial definition. The reason for our choice is that this definition has a straightforward extension to semi-cubical sets.

\begin{notation}[Sequences of integers]
  We denote sequences of integers by $\{i_1, \ldots i_n\}$. In particular, the empty sequence is written $\{\}$.
\end{notation}

@@

\begin{example}[$\Cube^1$]
  $\Cube^1$ consists of two points, given by $\Hom(0, 1)$ and a line, given by $\Hom(1, 1)$:

@@

\begin{example}[$\Cube^2$]
  $\Cube^2$ consists of four points, given by $\Hom(0, 2)$, four lines connecting the four points, given by $\Hom(1, 2)$, and a filler, given by $\Hom(2, 2)$:

@@

\begin{table}[H]
  \centering
  \begin{tabularx}{0.95\linewidth}{X|X|X}
    \toprule
    Value of $\nu$ & 1                               & 2                  \\
    \graymidrule
    Interpretation & Augmented semi-simplicial types & Semi-cubical types \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

In category theory, the correspondence between the fibered and indexed way manifests as the correspondence between Grothendieck fibrations and pseudofunctors, of which a particular instance is the correspondence between discrete fibrations and presheaves, or even between discrete fibrations over sets and families of sets.

@@

$\restrl[n][p][\epsilon](d)(c_L, c_R)$ produces a layer, in which the $\omega$-component has the type

@@

\subsection{Formal construction in ETT}
% Abbreviated tables in this section
%
% The rule for implicit arguments for each table:
% show 1 explicit in first two parts
% show E + 1 explicit in the third part
\appendmask[bonak]{layer}[D]
\appendmask[bonak]{filler}[D]
\appendmask[bonak]{restrframe}[D]
\appendmask[bonak]{restrlayer}[D, d]
\appendmask[bonak]{restrfiller}[D, d]
\appendmask[bonak]{cohframe}[D]
\appendmask[bonak]{cohlayer}[D, d]
\appendmask[bonak]{cohfiller}[D, d]

@@

\subsection{Formal construction in ITT\label{sec:itt}}
\renewcommandx{\cohf}{\coh{frame}[][][][][][][][]}

@@

\renewcommand*{\lab}{fulltab:coh}
\input{tab-coh.tex}
